---
title: "HR"
author: "Mike Pheasant"
date: "01/05/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
# install.packages("plotly")
# install.packages("corrplot")
library(plotly)
library(tidyverse)
library(corrplot)

```

## Load and Summarize Data

Derive a location (Brisbane, San Francisco or London) from the employee role:
* 90% of IT staff in Brisbane (5% in each other location)
* 90% sales in San Francisco (5% in each other location)
* 90% other categories in London (5% in each other location)

```{r}
location <- function(role) {
  prob=switch(as.character(role),
              IT=c(0.9, 0.05, 0.05),
              sales=c(0.05, 0.9, 0.05),
              c(0.05, 0.05, 0.9)
  )
  sample(c('Brisbane','San Francisco','London'), 1, rep=TRUE, prob=prob)
}
```

Load raw data (from https://www.kaggle.com/ludobenistant/hr-analytics), add factors, save, and summarise.

```{r}
HR <- read_csv("~/play/HR-kaggle/HR_comma_sep.csv", col_types=cols(
  satisfaction_level = col_double(),
  last_evaluation = col_double(),
  number_project = col_integer(),
  average_montly_hours = col_integer(),
  time_spend_company = col_integer(),
  Work_accident = col_integer(),
  left = col_integer(),
  promotion_last_5years = col_integer(),
  sales = col_character(),
  salary = col_character()
))

HR$salary <- factor(HR$salary, levels=c("low", "medium", "high"), ordered=TRUE)
HR$sales <- factor(HR$sales)
HR$location <- factor(sapply(HR$sales, location))

HR$Work_accident_ <- factor(HR$Work_accident == 1)
HR$promotion_last_5years_ <- factor(HR$promotion_last_5years == 1)
HR$left_ <- factor(HR$left==1)

write_csv(HR, "~/play/HR-kaggle/HR.csv")

summary(HR)
```

## Exploratory Data Analysis

### Numeric Correlations 

Leaving is negatively correlated with:

* satisfaction level (-0.39) 
* having a work accident (-0.15) 
* and less so with a promotion in the last 5 years (-0.06)

Leaving is positively correlated with:

* the time spent with the company (0.14)
* and less so with average hours worked per month (0.07)

In addition:

* satisfaction level is correlated with last evaluation (0.11), and negatively with number of projects (-0.14) and time spent with company (-0.10)
* last evaluation is positively correlated with number of projects (0.35), average monthly hours (0.34) and time spent with company (0.13)

```{r}
M <- cor(HR[1:8])
corrplot(M, method = 'circle', order ="hclust",
         tl.col="black", tl.cex = 1, tl.offset = 0.1, tl.srt = 45)
```

### Career advancement

Many people leaving have been with the company at least 4 years but have not had a promotion in the last 5 years, despite working long hours.

```{r}
with(HR, coplot(average_montly_hours ~ jitter(time_spend_company) |
                promotion_last_5years_ + left_))
```

These unpromoted, long-term workers who leave are receiving the highest evaluations.

```{r}
with(HR, coplot(last_evaluation ~ jitter(time_spend_company) |
                promotion_last_5years_ + left_))
```

These unpromoted, long-term workers who leave are also reporting high levels of satisfaction with the company.

```{r}
with(HR, coplot(satisfaction_level ~ jitter(time_spend_company) |
                promotion_last_5years_ + left_))
```

Of the unpromoted workers putting in long hours who are leaving, some report the lowest levels of satisfaction, but more report very high levels of satisfaction. 

The ones who are working the fewest hours report fairly low satisfaction. 

```{r}
with(HR, coplot(average_montly_hours ~ satisfaction_level |
                promotion_last_5years_ + left_))
```

Workers who put in many hours on many projects tend to leave if they havent had a promotion in the last 5 years.

```{r}
with(HR, coplot(average_montly_hours ~ jitter(number_project) |
                promotion_last_5years_ + left_))
```

Of the unpromoted workers who left, many had both high evaluations and high levels of satisfaction.

```{r}
with(HR, coplot(last_evaluation ~ satisfaction_level |
                promotion_last_5years_ + left_))
```




